---
format:
  pdf:
    engine: knitr
    geometry:
      - top=20mm
      - bottom=20mm
      - left=20mm
      - right=20mm
    include-in-header: 
      text: |
        \usepackage{fvextra}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
        \pagenumbering{gobble}
bibliography: utils/references.bib
---

```{r setup}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# {{ species }}

\vspace{-0.5cm}

<!-- *`r params$scientific_name`* -->

Overall vulnerability Rank =

Biological Sensitivity =

Climate Exposure =

Data Quality =

```{r,echo=FALSE, fig.align='left'}
here::here("test_images/{{ species }}_exposure_summary_table.png")
```

```{r,echo=FALSE, fig.align='left'}

here::here("test_images/{{ species }}_sensitivity_summary_table.png")

```

```{r,echo=FALSE, fig.align='left'}
here::here("test_images/{{ species }}_exposure_allvars.png")
```

```{r,echo=FALSE, fig.align='left'}
here::here("test_images/{{ species }}_exposure_impvars.png")
```

```{r,echo=FALSE, fig.align='left'}
here::here("test_images/{{ species }}_distribution_center_of_gravity.png")
```

```{r,echo=FALSE, fig.align='left'}
here::here("test_images/{{ species }}_distribution_area_of_high_probability.png")
```

```{r,echo=FALSE, out.width="200%", fig.align='left'}
here::here("test_images/{{ species }}_sdm.png")
```

\newpage

# {{ species }}

\vspace{-0.5cm}

<!-- *`r params$scientific_name`* -->

## Overall Climate Vulnerability Rank:

add text here

## Climate Exposure:

add text here

## Biological Sensitivity:

add text here

## Distributional Vulnerability Rank:

add text here

## Directional Effect:

add text here

## Data Quality:

add text here

## Climate Effects on Abundance and Distribution:

## Life History Synopsis:

## References
